\stktokens{}: Enforcing Well-Bracketed Control Flow and Stack Encapsulation Using Linear Capabilities

### 2019-07-27 popl-2019-journal-v/proving-full-abstraction.tex
[POPLjv] Change sum to cup

%%% spatial/spatial_owned
We cannot let multiple linear capabilities reference the same memory.
The spatial region represent ownership a part of memory which ensures that a linear capability is not aliased.
Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
To this end, we have shadow regions.
A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
This does not mean that the memory is necessarily not referenced.
A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t. ownership in Section~\ref{subsubsec:joining-worlds}).
Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \\ 
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
  \end{array} \right.
\]
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

% What parts of memory are linear
We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we continue the definitions of the three sub-worlds.
The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and normal capabilities, so it should contain both shared and spatial regions.
For this reason, it is defined as
\[
  \Worldh = \RegName \parfun (\Regionh + \Regions)
\]
\todo[inline]{Dominique: $+$?  Why not $\cup$?}
\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
For instance the $\Worldfs$ is defined as
\[
  \Worldfs = \RegName \parfun \Regions
\]
$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
Conceptually, this means that each of the stack frames is connected with a return point in some code.
In a traditional C calling convention, the code return point would even be stored in the stack frame.
However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
To this end, each shared region in $\Worlds$ must be paired with a return address:
\[
\Worlds = \RegName \parfun (\Regions \times \Addr)
\]

@@

Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities, in some sense, can be revoked.
A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory which makes it safe to repurpose.
On the other hand, normal capabilities may have an alias, so it is not safe to repurpose the memory a normal capability has authority over.
Only spatial regions specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
The $\revoked$ region does not specify any memory, so we drop the world indexed relation.
That is, we have
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) + \\
    \{\revoked \}
  \end{array} \right.
\]
The rules for the future region relation are displayed in Figure~\ref{fig:ft-reg-rel}.
% Revocation of spatial regions. 'revocation' of linear capabilities.
The first rule says that all regions are future regions of themselves, so regions are allowed to stay unchanged.
The second rule says that $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
The third and final rule says that $\spatial$ regions can become $\spatialo$ which indicates that the linear capabilities are affine.
Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes the linear capabilities affine.
The world a capability is safe with respect to is no longer needed when the capability is dropped, so other worlds may pick up the ownership which the future region relation allows.
% describe why "Spatial -> spatialo" ~ affine or at least why we have it?
\begin{figure}[htb]
  \centering
  \begin{mathpar}
  \inferrule{ r \in \Regions \cup \Regionh }{ r \future r}
  \and
  \inferrule{ }{ \revoked \future (\spatial,\_)}
  \and
  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
\end{mathpar}
  \caption{Future region relation.}
  \label{fig:ft-reg-rel}
\end{figure}

@@

The memory invariants are world dependent, but the worlds may change according to the future world relation, so what does that mean for the memory invariants?
The memory invariants describe what certain parts of memory looks like.
In future worlds, part of the memory may change but for the memory invariant to be invariant, it should continue relating the same memory segments even when the rest of the world changes.
\dominique{This discussion might be easier if we had talked about monotonicity earlier.}
For this reason, we require the world-indexed memory invariants to be monotone in the world.
This means that a pair of memory segments that are related now will stay related in future worlds.
This changes the spatial regions as follows:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) + \\
    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) + \\
    \{\revoked \}
  \end{array} \right.
\]
We make a similar change to $\Regionh$.
% TODO update terminology: memory invariant, H, and seal invariant, H_\sigma

@@

Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
The worlds we sketched have additional structure in terms of a future world relation.
The future world relation is captured by adding a preorder to c.o.f.e.'s:
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
We will need the standard construction for later, product, and sum later, so we define them here.
The reader may skip these definitions for now and refer back to them later.
\begin{definition}[$\blater$ preordered c.o.f.e.]
  \label{def:later-p-cofe}
  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
  \]
  where
  \[
    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{} \\
        x \nequal
        [n]_X x'
      \end{array}
    \right.
  \]
\end{definition}
We will also need the canonical preordered c.o.f.e.\ for sum.
\begin{definition}[Product preordered c.o.f.e.]
  \label{def:product-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
  \]
\end{definition}
\begin{definition}[Sum preordered c.o.f.e.]
  \label{def:sum-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X + Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
  \[
    \text{for } z,z' \in X + Y,\quad  z' \future z \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z' \future_x z \vee{} \\
        z,z' \in Y \wedge z' \future_Y z
      \end{array}
    \right.
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
    \text{for } z,z' \in X + Y,\quad  z \nequal z' \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z \nequal_x z' \vee{} \\
        z,z' \in Y \wedge z \nequal_Y z'
      \end{array}
    \right.
  \]
\end{definition}

@@

Next, we define the preordered c.o.f.e.\ for shared regions.
The future region relation on shared regions is non-trivial, so the definition of the preordered c.o.f.e.\ is a bit more involved.
\begin{definition}[Shared region preordered c.o.f.e.\ ]
  \label{def:shared-reg-p-cofe}
  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
  \[
    \Regions = \{\spatial\}\times \var{MemInv} + \{\spatialo\} \times \var{MemInv'} + \{\revoked\}
  \]
  define the preordered c.o.f.e.\ of shared regions as
  \[
  \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
  \]
  where the preorder $\future$ is defined as
  \begin{mathpar}
  \inferrule{ r \in \Regions }{ r \future r }
  \and
  \inferrule{ }{ \revoked \future (\spatial,\_)}
  \and
  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
  \end{mathpar}
  and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
  for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
  \begin{itemize}
  \item $n=0$
  \item $n > 0$ and one of the following is true
    \begin{itemize}
    \item $r = r' = \revoked$
    \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
    \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
    \end{itemize}
  \end{itemize}
\end{definition}
Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:shared-reg-p-cofe} define the preoreded c.o.f.e.'s necessary for the regions in our worlds.

@@

Finally, we define the sub-world for the heap.
To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the sum preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
The result is a preordered c.o.f.e.\
\[
  \Regions + \Regionh
\]
which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
\[
  \Worldh = \nats \parfun (\Regions + \Regionh)
\]

@@

\begin{theorem}
  \label{thm:rec-dom-eq-sol}
  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
  \[
    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
  \]
  and for $\hat{W},\hat{W}' \in \Wor$
  \[
    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
  \]
  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
\[
  \Worldh = \RegionName \parfun (\Regions + \Regionh)
\]
and
\[
  \Worlds = \RegionName \parfun (\Regions \times \Addr)
\]
and
\[
  \Worldfs = \RegionName \parfun \Regions
\]
where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
\end{multline*}
and
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) +\\
    \{\revoked\}
  \end{array} \right.
\]
\end{theorem}
% TODO informally describe how solution to recursive domain equation is found (in terms of what we have to show, not the fixed-point theorem).
%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
%% Aleš note? Tutorial note?
The solution to the recursive domain equation is the preordered c.o.f.e.\ $\Wor$ in Theorem~\ref{thm:rec-dom-eq-sol}.
Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
\[
  \World = \Worldh \times \Worlds \times \Worldfs
\]

>>>>

%%% spatial/spatial_owned
We cannot let multiple linear capabilities reference the same memory.
The spatial region represent ownership a part of memory which ensures that a linear capability is not aliased.
Even though spatial regions gives the right to reference part of memory, we still need the world to specify the remainder of the memory that may be referenced by linear capabilities.
To this end, we have shadow regions.
A shadow region is a shadow copy of a spatial region in the sense that it specifies part of memory, but it does not give the right to reference that part of memory.
This does not mean that the memory is necessarily not referenced.
A compatible world may claim a shadow region as spatial which allows the other world to reference the memory (we return to the notion of compatible worlds w.r.t. ownership in Section~\ref{subsubsec:joining-worlds}).
Specifically, we add tags $\spatialo$ and $\spatial$ to the spatial regions:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\ 
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg})  \\
  \end{array} \right.
\]
For readability, we also add a tag $\pure$ to the shared regions:
\[
  \Regionh = \{\pure \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) 
\]

% What parts of memory are linear
We will extend the regions further in Sections~\ref{subsubsec:seals}~and~\ref{subsubsec:ft-and-revocation}, but for now we continue the definitions of the three sub-worlds.
The sub-world $\Worldh$ specifies the heap memory which can be referenced by both linear and normal capabilities, so it should contain both shared and spatial regions.
For this reason, it is defined as
\[
  \Worldh = \RegName \parfun (\Regionh \cup \Regions)
\]
\srccm{} internalizes the \stktokens{} stack, so it can only be referenced by linear capabilities which means that the two stack regions should only have spatial regions.
For instance the $\Worldfs$ is defined as
\[
  \Worldfs = \RegName \parfun \Regions
\]
$\Worlds$ can also only be referenced by linear capabilities, so it should also only have spatial regions.
$\Worlds$ not only models the memory contents of the local stack frames, it also models the call stack that consists of the stack frames.
Conceptually, this means that each of the stack frames is connected with a return point in some code.
In a traditional C calling convention, the code return point would even be stored in the stack frame.
However, with \stktokens{} a stack frame does not contain any information about the corresponding code return point.
Instead, the stack frame and code return point is connected by the capabilities that reference them as they are sealed with the same seal and together then constitute the sealed return pair.
In order to ensure that each call actually returns to the correct point of the code, we must still include the address of the return point in our model.
To this end, each shared region in $\Worlds$ must be paired with a return address:
\[
\Worlds = \RegName \parfun (\Regions \times \Addr)
\]

@@

Specifically, it should only be the regions that specify memory governed by linear capabilities as linear capabilities, in some sense, can be revoked.
A linear capability is the sole authority over some memory, so when a program possesses a linear capability, it knows for sure that no one else has access to that memory which makes it safe to repurpose.
On the other hand, normal capabilities may have an alias, so it is not safe to repurpose the memory a normal capability has authority over.
Only spatial regions specify memory governed by linear capabilities, so we only add the revoked tag to the $\Regions$.
The $\revoked$ region does not specify any memory, so we drop the world indexed relation.
That is, we have
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\spatialo \} \times (\World \fun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\revoked \}
  \end{array} \right.
\]
The rules for the future region relation are displayed in Figure~\ref{fig:ft-reg-rel}.
% Revocation of spatial regions. 'revocation' of linear capabilities.
The first rule says that all regions are future regions of themselves, so regions are allowed to stay unchanged.
The second rule says that $\spatial$ regions can become $\revoked$ which means that nothing can depends on the region anymore.
The third and final rule says that $\spatial$ regions can become $\spatialo$ which indicates that the linear capabilities are affine.
Specifically, the linear capabilities in \trgcm{} and \srccm{} can be dropped by overwriting them in registers or memory which makes the linear capabilities affine.
The world a capability is safe with respect to is no longer needed when the capability is dropped, so other worlds may pick up the ownership which the future region relation allows.
% describe why "Spatial -> spatialo" ~ affine or at least why we have it?
\begin{figure}[htb]
  \centering
  \begin{mathpar}
  \inferrule{ r \in \Regions \cup \Regionh }{ r \future r}
  \and
  \inferrule{ }{ \revoked \future (\spatial,\_)}
  \and
  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
\end{mathpar}
  \caption{Future region relation.}
  \label{fig:ft-reg-rel}
\end{figure}

@@

The memory invariants are world dependent, but the worlds may change according to the future world relation, so what does that mean for the memory invariants?
The memory invariants describe what certain parts of memory looks like.
In future worlds, part of the memory may change but for the memory invariant to be invariant, it should continue relating the same memory segments even when the rest of the world changes.
\dominique{This discussion might be easier if we had talked about monotonicity earlier.}
For this reason, we require the world-indexed memory invariants to be monotone in the world.
This means that a pair of memory segments that are related now will stay related in future worlds.
This changes the spatial regions as follows:
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\spatialo \} \times (\World \monfun \Rel{\MemSeg \times \MemSeg}) \cup \\
    \{\revoked \}
  \end{array} \right.
\]
We make a similar change to $\Regionh$.
% TODO update terminology: memory invariant, H, and seal invariant, H_\sigma

@@

Unfortunately, the standard c.o.f.e.\ constructions do not suffice for the worlds we want to define.
The worlds we sketched have additional structure in terms of a future world relation.
The future world relation is captured by adding a preorder to c.o.f.e.'s:
\begin{definition}[Preordered c.o.f.e.]
  A preordered c.o.f.e.\ is a c.o.f.e.\ equipped with a preorder on $X$, $\left(X, \left(\nequal\right)_{n=0}^{\infty}, \future \right)$. 
  \begin{itemize}
  \item The ordering preserves limits. That is, for Cauchy sequences $\{a_n\}_n$ and $\{b_n\}_n$ in $X$ if $\{a_n\}_n \future\{b_n\}_n$, then $\lim \{a_n\}_n \future \lim \{b_n\}_n$.
  \end{itemize}
\end{definition}
The preorder must respect the existing structure of the c.o.f.e.'s which means that it must preserve limits.
There are also canonical constructions for preordered c.o.f.e.'s that simply rely on the preorder of the underlying preordered c.o.f.e.'s.
We will need the standard construction for later, product, and sum, so we define them here.
The reader may skip these definitions for now and refer back to them later.
\begin{definition}[$\blater$ preordered c.o.f.e.]
  \label{def:later-p-cofe}
  Given a preordered c.o.f.e $\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right)$ define
  \[
    \blater\left(X, \left(\nequal_X\right)_{n=0}^{\infty}, \future_X \right) = \left(X, \left(\nequal[n]_{\blater}\right)_{n=0}^{\infty}, \future_X \right)
  \]
  where
  \[
    x \nequal[n+1]_{\blater} x' \text{ iff } \left\{
      \begin{array}{l}
        n = 0 \vee{} \\
        x \nequal
        [n]_X x'
      \end{array}
    \right.
  \]
\end{definition}
We will also need the canonical preordered c.o.f.e.\ for sum.
\begin{definition}[Product preordered c.o.f.e.]
  \label{def:product-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_{Y}\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X \times Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x',y') \future (x,y) \text{ iff } x' \future_X x \wedge y' \future_Y y
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
\text{for } (x,y),(x',y') \in X \times Y, \quad    (x,y) \nequal (x',y') \text{ iff } x \nequal_X x' \wedge y \nequal_Y y'
  \]
\end{definition}
\begin{definition}[Union preordered c.o.f.e.]
  \label{def:sum-p-cofe}
  Given two c.o.f.e.'s $\left(X,\left(\nequal[n]_X\right)_{n=0}^{\infty}, \future_X \right)$ and $\left(Y,\left(\nequal[n]_Y\right)_{n=0}^{\infty}, \future_y \right)$, define the product preordered c.o.f.e.\ as
  \[
  \left(X \cup Y ,\left(\nequal[n]\right)_{n=0}^{\infty},\future \right)
  \]
  where the preorder $\future$ distributes to the underlying preorder, i.e.\ 
  \[
    \text{for } z,z' \in X \cup Y,\quad  z' \future z \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z' \future_x z \vee{} \\
        z,z' \in Y \wedge z' \future_Y z
      \end{array}
    \right.
  \]
  and the family of equivalences distributes to the underlying families of equivalences, i.e.\
  \[
    \text{for } z,z' \in X \cup Y,\quad  z \nequal z' \text{ iff } \left\{
      \begin{array}{l}
        z,z' \in X \wedge z \nequal_x z' \vee{} \\
        z,z' \in Y \wedge z \nequal_Y z'
      \end{array}
    \right.
  \]
\end{definition}

@@

Next, we define the preordered c.o.f.e.\ for shared regions.
The future region relation on shared regions is non-trivial, so the definition of the preordered c.o.f.e.\ is a bit more involved.
\begin{definition}[Shared region preordered c.o.f.e.\ ]
  \label{def:shared-reg-p-cofe}
  Given two c.o.f.e.'s $\left(\var{MemInv},\left(\nequal[n]_\var{MI}\right)_{n=0}^{\infty}\right)$ and $\left(\var{MemInv'},\left(\nequal[n]_\var{MI'}\right)_{n=0}^{\infty}\right)$, let
  \[
    \Regions = \{\spatial\}\times \var{MemInv} \cup \{\spatialo\} \times \var{MemInv'} \cup \{\revoked\}
  \]
  define the preordered c.o.f.e.\ of shared regions as
  \[
  \left(\Regions,\left(\nequal[n]\right)_{n=0}^{\infty},\future\right)
  \]
  where the preorder $\future$ is defined as
  \begin{mathpar}
  \inferrule{ r \in \Regions }{ r \future r }
  \and
  \inferrule{ }{ \revoked \future (\spatial,\_)}
  \and
  \inferrule{ }{ (\spatialo,H) \future (\spatial,H)}
  \end{mathpar}
  and $\left(\nequal[n]\right)_{n=0}^{\infty}$ is the total relation for $n=0$ and otherwise the region tags must be the same and the memory invariants $n$-equal, i.e.\ 
  for $r ,r' \in \Regions$ we have $r \nequal r'$ when one of the following holds
  \begin{itemize}
  \item $n=0$
  \item $n > 0$ and one of the following is true
    \begin{itemize}
    \item $r = r' = \revoked$
    \item $r = (\spatial,H)$, $r' = (\spatial,H')$, and $H \nequal_\var{MI} H'$
    \item $r = (\spatialo,H)$, $r' = (\spatialo,H')$, and $H \nequal_\var{MI'} H'$
    \end{itemize}
  \end{itemize}
\end{definition}
Definitions~\ref{def:pure-reg-p-cofe} and~\ref{def:shared-reg-p-cofe} define the preoreded c.o.f.e.'s necessary for the regions in our worlds.

@@

Finally, we define the sub-world for the heap.
To this end, we combine the shared region preordered c.o.f.e.\ (Definition~\ref{def:pure-reg-p-cofe}) with the shared region preordered c.o.f.e.\ (Definition~\ref{def:shared-reg-p-cofe}) using the union preordered c.o.f.e.\ construction in Definition~\ref{def:sum-p-cofe}.
The result is a preordered c.o.f.e.\
\[
  \Regions \cup \Regionh
\]
which we use in the sub-world construction (Definition~\ref{def:sub-world-p-cofe}) to obtain the preordered c.o.f.e.\
\[
  \Worldh = \nats \parfun (\Regions \cup \Regionh)
\]

@@

\begin{theorem}
  \label{thm:rec-dom-eq-sol}
  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e., and there exists an isomorphism $\xi$ such that
  \[
    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
  \]
  and for $\hat{W},\hat{W}' \in \Wor$
  \[
    \hat{W}' \future \hat{W} \text{ iff } \xi(\hat{W}') \future \xi(\hat{W})
  \]
  for $\Worlds$, $\Worldh$, and $\Worldfs$ defined as follows
\[
  \Worldh = \RegionName \parfun (\Regions \cup \Regionh)
\]
and
\[
  \Worlds = \RegionName \parfun (\Regions \times \Addr)
\]
and
\[
  \Worldfs = \RegionName \parfun \Regions
\]
where $\RegionName = \nats$. $\Regions$ and $\Regionh$ defined as follows
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps\times\SealableCaps})
\end{multline*}
and
\[
  \Regions = \left\{
  \begin{array}{l}
    \{\spatial, \spatialo\} \times (\Wor \monnefun \URel{\MemSeg\times\MemSeg}) \cup\\
    \{\revoked\}
  \end{array} \right.
\]
\end{theorem}
% TODO informally describe how solution to recursive domain equation is found (in terms of what we have to show, not the fixed-point theorem).
%% The category-theoretic solution of recursive metric-space equations, Birkedal et al?
%% Aleš note? Tutorial note?
The solution to the recursive domain equation is the preordered c.o.f.e.\ $\Wor$ in Theorem~\ref{thm:rec-dom-eq-sol}.
Note, that $\Wor$ is not equal to $\blater (\Worldh \times \Worlds \times \Worldfs)$; it is isomorphic.
This means that whenever we encounter $\Wor$, we have to apply $\xi$ and go under a later before we can actually use the world.
This makes it rather inconvenient to have $\Wor$ as the world, so instead we define the worlds as 
\[
  \World = \Worldh \times \Worlds \times \Worldfs
\]
